---
title: "Linear Regression"
format: pdf
editor: visual
---

## Module 3 - Activity 4

### Linear Regression

Of the models with a convex representation of their parametricestimation, generalized linear models (GLM) are a crucial case.The two most frequent examples of GLM are linear regression andlogistic regression. Therefore, in this activity, several linear regressionand logistic regression exercises will be solved with R software andsome of its packages.

#### Activities Problem 1: Warm Up

##### 1. Section 3.7 Problem 8.

This question involves the use of simple linear regression on the Auto data set.

##### a) Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output.

Adding the necessary libraries

```{r}
#| warning: false
library(ISLR2)
library(tidymodels)
library(dplyr)
```

Loading dataframe

```{r}
auto <- Auto
head(auto)
```

The Auto df counts with 9 columns that show information of different models of vehicles. For the present exercise is intended to evaluate if exists any relationship

```{r}
linearmodelauto <- lm(mpg ~ horsepower, data = auto)
linearmodelauto
```

-   Is there a relationship between the predictor and the response?

    Yes there is a relationship between the variables

```{r}
summary(linearmodelauto)
```

-   How strong is the relationship between the predictor and the response

    With the p values of the model (\<0.0001) we can assume that the model is significant to explain the relationship between our variables . Also the $R^{2}$ of 60% indicates that there is a highly correlation between this variables

-   Is the relationship between the predictor and the response positive or negative?

    The relationship is negative

-   What is the predicted mpg associated with a horsepower of 98? What are the associated 95 % confidence and prediction intervals?

```{r}
predict(linearmodelauto, tibble(horsepower=98), interval = "confidence")
```

##### 

b)  Plot the response and the predictor. Use the abline() functionto display the least squares regression line.

```{r}
auto %>%
  ggplot(aes(x = horsepower)) +
  geom_point(aes(y = mpg), size = 2, alpha = 0.4) +
  geom_abline(slope = coef(linearmodelauto)["horsepower"],
              intercept = coef(linearmodelauto)["(Intercept)"],
              size = 2, color = "blue")
```

##### c) Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.

```{r}
linearmodelauto %>% performance::check_model()
```

#####  
